---
title: "COVID-19 vaccine study Final2021"
author: "Sasha Sanchez"
date: "Summer 2021"
output: 
    html_document:
          toc: true
          toc_depth: 5
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor) )
guy = filter(FinalData, sex =="guy")
gal = filter(FinalData, sex =="gal")
gay = filter(FinalData, LGBTQ == "gay")
Druguser = filter(FinalData, DrugUser == "yes")
```


# Introduction

A new  preventative vaccine (COVID-19) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of sub populations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbiage when referencing p-values in conclusions.


# Guys

### Barplot

```{r}
dd2 <- guy %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This Bar plot represents the population of Guys meaning (Males) people and the relationship it has contraction of COVID-19 by taking the vaccine or the placebo.The left one shows the ones that got COVID and the right side shows the guys that didn't get COVID. 

The Bar plot above also shows that about 58% did shockingly get COVID-19 even though they got the vaccine. And about 42% of the placebo did in fact contracted COVID-19. There has been a high increase percentage on those who took the vaccine and contracted COVID-19 versus those who only took the placebo. It appears that the COVID-19 vaccine is effective by it's protection. The men that took the drug were one in half more than likely to get COVID-19 than those who took the placebo.

### Numerical Results


```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```


The numerical results for Guys (Males) that got COVID-19 after taking the placebo and taking the vaccine. Among the guys got a 58% got COVID-19 after taking the vaccine while the other half which is 42% of guys that were given the placebo contracted COVID-19. The COVID-19 vaccine is not effective in preventing COVID-19 in Guys. The odds on getting COVID-19 while taking the vaccine is 3:2.



### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The information above is very Strategic and can change. The p -value displays to be small going at a 0.0006 which means that we can reject the null hypothesis in favor of the alternative. The odd ratio is at a 1.5 which would explain the deaths of guys. 



```{r}
effectiveness =  1.4  /( 1.4   +1)
print(effectiveness*100)
```

```{r}
low=(1.7/(1.7 + 1))*100
hi=(1.8/(1.8 +1))*100
options(digits=2)
```




## Effectiveness Confidence Interval `r low`  to `r hi`


```{r}
chisq.test(table1)
chisqtestGC(table1)
malesdrug <- fisher.test(table1)
malesdrug
```


```{r}
effect <- malesdrug$p.value/(malesdrug$p.value +1)
effectivenessmales =malesdrug$estimate/(malesdrug$estimate +1)*100
print(effectivenessmales)
```


```{r}
low=(malesdrug$conf.int[1]/(malesdrug$conf.int[1] + 1))*100
hi=(malesdrug$conf.int[2]/(malesdrug$conf.int[2] + 1))*100
```



# Gals

### Barplot


```{r}
dd2 <- gal %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


In this bar plot shows the population of Gals meaning (Females) and the relationship of those who got the vaccine and those who got the placebo. 

In this plot it shows that those who were given the placebo were more contracted to COVID-19. Only about 33% who where given the vaccine resulted in contracting COVID-19. This just means that about 65% of those that got COVID-19 were the only ones who recieved the placebo. 

  
### Numerical Results  
  
  
```{r}
table1 <- xtabs(~infected + treatment, data=gal)
rowPerc(table1)
colPerc(table1)
```


The numerical results for the gals (Females) after getting the COVID-19 vaccine is at a 33% and a 67% for the placebo.The vaccine is effective in protecting gals against the contraction of COVID-19. This concludes the odds of 1:2 in getting COVID-19 while taking the vaccine. The gals who got the vaccine were more likely to get COVID-19 than those who took the placebo. 


### Inferntial Results


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The inferential results for the gals (Females) the p -value is super small it is stated as a 0.0000000000002. So therefore we are going to reject the null hypothesis. The odds are 0.49 as the given ratio displayed on the table which means the gals were 0.49 more likely to get COVID-19 than those who took the placebo.  


```{r}
effectiveness =  0.49  /( 0.49   +1)
print(effectiveness*100)
```

```{r}
low=(0.41/(0.58 + 1))*100
hi=(1.8/(1.8 +1))*100
options(digits=2)
```


```{r}
chisq.test(table1)
chisqtestGC(table1)
galsdrug <- fisher.test(table1)
galsdrug
```




```{r}
effect <- galsdrug$p.value/(galsdrug$p.value +1)
effectivenessgals =galsdrug$estimate/(galsdrug$estimate +1)*100
print(effectivenessgals)
```


```{r}
low=(galsdrug$conf.int[1]/(galsdrug$conf.int[1] + 1))*100
hi=(galsdrug$conf.int[2]/(galsdrug$conf.int[2] + 1))*100
```




## Effectiveness Confidence Interval `r low`  to `r hi`



# Gay Sexual Orientaion (LGBTQ)

### Barplot 

```{r}
dd2 <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This bar plot represents the population of Gay people meaning (LGBTQ) Community and the relationship of the contraction of COVID-19 by taking the vaccine or the placebo.

The bar plot above shows that about 71% of individuals who got the vaccine got COVID-19 followed by the 29% of the individuals who received the placebo. The individuals who received the vaccine were about three times more likely to get COVID-19 than those who took the placebo.

### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```


The numerical results shows that 71% of Gay individuals after taking the COVID-19 vaccine. And 29% who took the placebo.

### Inferential Results


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The inferential results show that the p -value is small which we are going to reject the null hypothesis. The odds of this is standing at a ratio of 2:5.



```{r}
effectiveness =  2.5  /( 2.5   +1)
print(effectiveness*100)
```

```{r}
low=(1.6/(3.9 + 1))*100
hi=(1.8/(1.8 +1))*100
options(digits=2)
```


```{r}
chisq.test(table1)
chisqtestGC(table1)
LGBTQdrug <- fisher.test(table1)
LGBTQdrug
```



```{r}
effect <- LGBTQdrug$p.value/(LGBTQdrug$p.value +1)
effectivenessLGBTQ =LGBTQdrug$estimate/(LGBTQdrug$estimate +1)*100
print(effectivenessLGBTQ)
```


```{r}
low=(LGBTQdrug$conf.int[1]/(LGBTQdrug$conf.int[1] + 1))*100
hi=(LGBTQdrug$conf.int[2]/(LGBTQdrug$conf.int[2] + 1))*100
```




## Effectiveness Confidence Interval `r low`  to `r hi`



# Drug Users 

### Barplot 



```{r}
dd2 <- Druguser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


In this Bar plot represents the population of Drug users who answered yes and the relationship to the contraction of COVID-19 while taking the vaccine or the placebo. 

The bar plot shows that about 34% who answered yes to the drug use and received the vaccine contracted COVID-19. This means that about 66% of the drug users that only took the placebo resulted in them getting COVID-19.  

### Numerical Results


```{r}
table1 <- xtabs(~infected + treatment, data=Druguser)
rowPerc(table1)
colPerc(table1)
```


The numerical results show that the Drug users who took the COVID-19 vaccine are at a 33% and the Drug users who took the placebo are at a 67%. With that being said the Gay individuals that got COVID-19 after taking the placebo and taking the vaccine contracted about 67%. The ones that got the placebo only contracted 33%. It became ineffectively in protecting those individuals. The ones who took the vaccine were about three times more likely to get COVID-19 than those who took the placebo. 

### Inferntial Results


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The inferential results display that the p -value is at 0.002 which is very small. So we can reject the null hypothesis. The odds of the ratio is at 0.55 so that means that the drug is effective in preventing the chances of getting COVID-19 in Drug users.



```{r}
effectiveness =  0.5  /( 0.5   +1)
print(effectiveness*100)
```

```{r}
low=(0.36/(0.82 + 1))*100
hi=(1.8/(1.8 +1))*100
options(digits=2)
```


```{r}
chisq.test(table1)
chisqtestGC(table1)
drugfish <- fisher.test(table1)
drugfish
```


```{r}
effect <- drugfish$p.value/(drugfish$p.value +1)
effectivenessdrug =drugfish$estimate/(drugfish$estimate +1)*100
print(effectivenessdrug)
```


```{r}
low=(drugfish$conf.int[1]/(drugfish$conf.int[1] + 1))*100
hi=(drugfish$conf.int[2]/(drugfish$conf.int[2] + 1))*100
```




## Effectiveness Confidence Interval `r low`  to `r hi`



# Conclusion And Recommendations 

Generally speaking about the COVID-19 vaccine, It appeared that in some cases the vaccine had an affect and so did the placebo among different Subjects. The vaccine was not effective to the Males However, Females were protected by COVID-19. The Covid vaccine was not effective for those who were in the LGBTQ Community. As for the Drug users they were fully protected by the vaccine.Our best recommendation is that if you are a Male or someone who identifies with being Gay it is not recommended for one to get the vaccine. On the other hand, if you are a Female or a Drug user the vaccine came about being very effective. The COVID-19 effectiveness for the Males was a 62.96 which indicates that the vaccine was not effective in protecting men contracting diseases. The COVID-19 vaccine was successful on Females by protecting them against COVID-19 which stated as 29.01. The COVID-19 vaccine was not effective at all for those individuals that are part of the LGBTQ community leaving the number to a 62.03. And lastly the Drug users were protected by the vaccine indicating a 26.37.  
